---
title: "Final Exam: Trends in Opioid Misuse and Abuse in the United States Between 2016 and 2017"
author: "Bernadette F. Burks, Department of Graduate and Professional Studies, New England College - CRN227: Information Visualization, Prof. Alexander Stodolski"
date: "July 2, 2024"
output:
  word_document: default
  html_document: default
---

# "Predicting Pain Reliever Misuse and Abuse"  


**Introduction to Code**  


I have imported and utilized the following datasets for this assignment:

Visualizations 1-4:  
https://www.kaggle.com/datasets/thedevastator/predicting-pain-reliever-misuse-abuse/code?datasetId=2837566/

Visualization 5 (Map):
https://www.cdc.gov/nchs/pressroom/sosmap/drug_poisoning_mortality/drug_poisoning.htm

***  


**About this File**  

This dataset contains information on demographics, medication use, and illicit drug use to predict the risk of pain reliever misuse and abuse.  


_Variable Names & Features_  

PRLMISAB  
Prescription opioid pain reliever misuse and abuse (Likert scale: 0-12)  

YEAR  
Year of NSDUH survey (15=2015, 16=2016, 17=2017)   

AGECAT  
Age category (1=12-17 years, 2=18-25, 3=26-35, 4=36-49, 5=50 and older)   

SEX  
Sex (0=Male, 1=Female)  

MARRIED  
Marital status (0=unmarried, 1=divorced, 2=widowed, 3=married)   

EDUCAT  
Education level (1=h.s. or Less, 2=h.s. grad., 3=some college, 4=college grad.)   

EMPLOY18  
Employment status, over age 18 (1=not employed, 2=part-time, 3=full-time)   

CTYMETRO  
Size of city/metropolitan region (1=rural, 2=small, 3=large)   

HEALTH  
Health problems, aggregated (Likert scale: 0-10)   

MENTHLTH  
Mental health, aggregated: adult depression, emotional distress (Likert scale: 0-10)   

TRTMENT  
Treatment for drugs or alcohol in past year, aggregated (Likert scale: 0-5)   

MHTRTMT  
Mental health treatment, aggregated (Likert scale: 1-10)   

TRQLZRS  
Tranquilizer use in past year, aggregated (Likert scale: 0-5)   

SEDATVS  
Sedative use in past year, aggregated (Likert scale: 0-5)   

HEROINUSE  
Heroin use in past year, aggregated (Likert scale: 0-5)   

COCAINE  
Cocaine and crack cocaine use in past year, aggregated (Likert scale: 0-5)   

AMPHETMN  
Amphetamine and methamphetamine use in past year, aggregated (Likert scale: 0-5)   



Source: https://doi.org/10.5281/zenodo.2301844  


***

**Package and Library Setup**
```{r message=FALSE, warning=FALSE}
# install.packages("tidyverse", "ggplot2", "ggthemes", "readr")

library(tidyverse) #loads library for pipe operation setup
library(ggplot2) #loads library for graphic visualization features
library(ggthemes) #loads pre-built themes for use with ggplot2
library(readr) #loads library to import datasets into R studio environment
```


***

**Load Raw Data**

```{r}
painrl <- read_csv("prlmis-data-full.csv", show_col_types = FALSE) #uses readr to import my primary dataset, naming it painrl (short for "pain relievers")
#summary(painrl) - commented out but shows an overview of the data's structure and variables
```
***

**Mutate Data for Usability (for string variables)**

```{r}
#Year of NSDUH survey (15=2015, 16=2016, 17=2017) is called YEAR
painrl <- painrl %>% #pipe operator setup
  mutate(YEAR = str_replace(YEAR, "15", "2015")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(YEAR = str_replace(YEAR, "16", "2016")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(YEAR = str_replace(YEAR, "17", "2017")) #mutate function to update value in dataset

#Age category (1=12-17 years, 2=18-25, 3=26-35, 4=36-49, 5=50 and older) is called AGECAT
painrl <- painrl %>% #pipe operator setup
  mutate(AGECAT = str_replace(AGECAT, "1", "2-17 years")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(AGECAT = str_replace(AGECAT, "2", "18-25 years")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(AGECAT = str_replace(AGECAT, "3", "26-35 years")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(AGECAT = str_replace(AGECAT, "4", "36-49 years")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(AGECAT = str_replace(AGECAT, "5", "50 and older")) #mutate function to update value in dataset

#Sex (0=Male, 1=Female) is called SEX
painrl <- painrl %>% #pipe operator setup
  mutate(SEX = str_replace(SEX, "0", "Male")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(SEX = str_replace(SEX, "1", "Female")) #mutate function to update value in dataset

#Marital status (0=unmarried, 1=divorced, 2=widowed, 3=married) is called MARRIED
painrl <- painrl %>% #pipe operator setup
  mutate(MARRIED = str_replace(MARRIED, "0", "Single"))  #mutate function to update value in dataset  Note: changed "unmarried" to "single" to remove biased language (ie - "childless" sounds biased, "child-free" does not).
painrl <- painrl %>% #pipe operator setup
  mutate(MARRIED = str_replace(MARRIED, "1", "Divorced")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(MARRIED = str_replace(MARRIED, "2", "Widowed")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(MARRIED = str_replace(MARRIED, "3", "Married")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(MARRIED = str_replace(MARRIED, "4", "NA"))  #mutate function to update value in dataset

#Education level (1=h.s. or Less, 2=h.s. grad., 3=some college, 4=college grad.) is called EDUCAT
painrl <- painrl %>% #pipe operator setup
  mutate(EDUCAT = str_replace(EDUCAT, "1", "Some High School")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(EDUCAT = str_replace(EDUCAT, "2", "High School Graduate")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(EDUCAT = str_replace(EDUCAT, "3", "Some College")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(EDUCAT = str_replace(EDUCAT, "4", "College Graduate")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(EDUCAT = str_replace(EDUCAT, "5", "NA")) #mutate function to update value in dataset


#Employment status, over age 18 (1=not employed, 2=part-time, 3=full-time) is called EMPLOY18
painrl <- painrl %>% #pipe operator setup
  mutate(EMPLOY18 = str_replace(EMPLOY18, "1", "Not Currently Employed")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(EMPLOY18 = str_replace(EMPLOY18, "2", "Employed Part-Time")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(EMPLOY18 = str_replace(EMPLOY18, "3", "Employed Full-Time")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(EMPLOY18 = str_replace(EMPLOY18, "0", "NA")) #mutate function to update value in dataset

#Size of city/metropolitan region (1=rural, 2=small, 3=large) is called CTYMETRO
painrl <- painrl %>% #pipe operator setup
  mutate(CTYMETRO = str_replace(CTYMETRO, "1", "Rural Area")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(CTYMETRO = str_replace(CTYMETRO, "2", "Suburban")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(CTYMETRO = str_replace(CTYMETRO, "3", "Large City")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(CTYMETRO = str_replace(CTYMETRO, "0", "NA")) #mutate function to update value in dataset
```
***

**Mutate Data for Usability (for all "Likert Scale" variables)**

```{r}

#heroine use in past year, on a scale from 0 - 5
painrl <- painrl %>% #pipe operator setup
  mutate(HEROINUSE = str_replace(HEROINUSE, "0", "Very unlikely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(HEROINUSE = str_replace(HEROINUSE, "1", "Unlikely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(HEROINUSE = str_replace(HEROINUSE, "2", "Somewhat unlikely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(HEROINUSE = str_replace(HEROINUSE, "3", "Somewhat likely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(HEROINUSE = str_replace(HEROINUSE, "4", "Likely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(HEROINUSE = str_replace(HEROINUSE, "5", "Very Likely")) #mutate function to update value in dataset


#TRQLZRS use in past year, on a scale from 0 - 5
painrl <- painrl %>% #pipe operator setup
  mutate(TRQLZRS = str_replace(TRQLZRS, "0", "Very unlikely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(TRQLZRS = str_replace(TRQLZRS, "1", "Unlikely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(TRQLZRS = str_replace(TRQLZRS, "2", "Somewhat unlikely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(TRQLZRS = str_replace(TRQLZRS, "3", "Somewhat likely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(TRQLZRS = str_replace(TRQLZRS, "4", "Likely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(TRQLZRS = str_replace(TRQLZRS, "5", "Very Likely")) #mutate function to update value in dataset

#SEDATVS use in past year, on a scale from 0 - 5
painrl <- painrl %>% #pipe operator setup
  mutate(SEDATVS = str_replace(SEDATVS, "0", "Very unlikely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(SEDATVS = str_replace(SEDATVS, "1", "Unlikely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(SEDATVS = str_replace(SEDATVS, "2", "Somewhat unlikely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(SEDATVS = str_replace(SEDATVS, "3", "Somewhat likely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(SEDATVS = str_replace(SEDATVS, "4", "Likely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(SEDATVS = str_replace(SEDATVS, "5", "Very Likely")) #mutate function to update value in dataset

#COCAINE use in past year, on a scale from 0 - 5
painrl <- painrl %>% #pipe operator setup
  mutate(COCAINE = str_replace(COCAINE, "0", "Very unlikely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(COCAINE = str_replace(COCAINE, "1", "Unlikely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(COCAINE = str_replace(COCAINE, "2", "Somewhat unlikely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(COCAINE = str_replace(COCAINE, "3", "Somewhat likely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(COCAINE = str_replace(COCAINE, "4", "Likely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(COCAINE = str_replace(COCAINE, "5", "Very Likely")) #mutate function to update value in dataset

#AMPHETMN use in past year, on a scale from 0 - 5
painrl <- painrl %>% #pipe operator setup
  mutate(AMPHETMN = str_replace(AMPHETMN, "0", "Very unlikely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(AMPHETMN = str_replace(AMPHETMN, "1", "Unlikely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(AMPHETMN = str_replace(AMPHETMN, "2", "Somewhat unlikely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(AMPHETMN = str_replace(AMPHETMN, "3", "Somewhat likely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(AMPHETMN = str_replace(AMPHETMN, "4", "Likely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(AMPHETMN = str_replace(AMPHETMN, "5", "Very Likely")) #mutate function to update value in dataset

#HALUCNG use in past year, on a scale from 0 - 5
painrl <- painrl %>% #pipe operator setup
  mutate(HALUCNG = str_replace(HALUCNG, "0", "Very unlikely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(HALUCNG = str_replace(HALUCNG, "1", "Unlikely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(HALUCNG = str_replace(HALUCNG, "2", "Somewhat unlikely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(HALUCNG = str_replace(HALUCNG, "3", "Somewhat likely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(HALUCNG = str_replace(HALUCNG, "4", "Likely")) #mutate function to update value in dataset
painrl <- painrl %>% #pipe operator setup
  mutate(HALUCNG = str_replace(HALUCNG, "5", "Very Likely")) #mutate function to update value in dataset


summary(painrl) #reprints our clean, updated dataset with more "readable" value strings
```
***


**Dropping NA Character Values From Set**
```{r}
painrl_new <- painrl[!grepl('NA', painrl$MARRIED),] #updates dataset by dropping NA values
painrl_new2 <- painrl_new[!grepl('NA', painrl_new$EDUCAT),] #updates dataset by dropping NA values
painrl_new3 <- painrl_new2[!grepl('NA', painrl_new2$EMPLOY18),] #updates dataset by dropping NA values
painrl_new_final <- painrl_new3[!grepl('NA', painrl_new3$CTYMETRO),] #updates dataset by dropping NA values
```

The "painrl_new_final" dataset is clean and removed of all rows with at least 1 or more missing/null values. Most of the demographic values have been updated from binary/integers to characters for readability and for experimentation (other analyses of this dataset have utilized only numeric values).  


***

**Final Details of Our Clean Dataset**  

```{r}
summary(painrl_new_final) #we can see that we ended up with 25,015 individual survey participants to use for our visualizations. This is still a very large set, so we are good to move forward on constructing some visualizations!
```


***

**Visualizations**  

## Visualization #1 - Self-Ranked Mental Health Status by Type of Populated Area   

Hypothesis/Question Presented - How do individuals in populous vs rural areas rank their own mental health on a scale from 0-10, with 0 being the worst and 10 being the best/highest rating?  

```{r warning=FALSE}
library(tidyverse) #loads the tidyverse library for pipe operator use

painrl_new_final %>% #sets up pipe operation
  ggplot(aes(x = MENTHLTH)) + #creates ggplot visualization with mental health rating on the x-axis
  geom_histogram(fill = "orange", binwidth = 4)+ #overlays the histogram setting with orange fill and a binwidth of 4
  theme_hc()+ #a pre-built theme via the ggthemes library
  scale_x_continuous(limits = c(0, 10), breaks = seq(0, 10, by = 1))+ #sets the x-axis integers/breaks
  scale_y_continuous(limits = c(0, 2000), breaks = seq(0, 2000, by =500))+ #sets the y-axis integers/breaks
  facet_grid(~CTYMETRO)+ #splits our 3 "CTYMETRO" variables into separate graphs side by side
  labs(x="Scale from 0 - 12", y="Individuals by Population",title="Self-Ranked Mental Health Status by Type of Populated Area", subtitle = "Reported via Likert 10-Point Scale | 0 = Worst and 12 = Best") #labels the graph with title and subtitle, x-axis, and y-axis
```

Analysis/Findings...  This visualization shows us that the general "trend" and shape of the data is relatively similar across the board. We can see that, in all types of populated areas, individuals in the majority rank their own mental health on a scale of 0-10 at around 2-6. There does appear to be a more equal distribution of individuals in rural areas on the 0-10 mental health scale - the majority in rural areas who feel their mental health is only 2-6 range is less severe than in suburban and urban landscapes. We can say, based on this data alone, that there may be a slightly higher overall mental health and well being level of individuals per capita in rural areas versus their higher populated counterparts.


***


## Visualization #2 - The Effects of Sex & Employment Status on Opioid Abuse Potential  

Hypothesis/Question Presented - What effect do both sex (male/female) and employment status (not employed/employed) have on an individual's self-rating of how likely they are to abuse prescription opioids, on a Likert scale of 0-12, with 0 = Would not abuse and 12 = Extremely Likely to Abuse?

```{r warning=FALSE}
ggplot(painrl_new_final, aes(x = PRLMISAB, fill = SEX)) + #creates ggplot visualization using the painrl_new_final dataset with PRLMISAB rating (0-12 Likert scale of desire to abuse) on the x-axis and sex on the y-axis
  geom_bar(position = "dodge", width = 3) + #overlays a bar plot setting with a stacked view of the data
  facet_wrap(~EMPLOY18) + #cuts apart our data groups by sex
  labs(title = "The Effects of Sex & Employment Status on Opioid Abuse Potential", subtitle = "Likert 12-Point Scale with | 0 = Would Not Abuse and | 12 = Extremely Likely to Abuse", y = "Number of Individuals", x = "Likert Scale Between 0-12", fill = "Sex") + #labels the graph with title and subtitle, x-axis, and y-axis
  theme_few()+ #a minimalist, pre-built theme from ggthemes
  scale_x_continuous(limits = c(0, 12), breaks = seq(0, 12, by = 6))+ #sets the x-axis integers/breaks
  scale_y_continuous(limits = c(0, 800), breaks = seq(0, 800, by = 100)) #sets the y-axis integers/breaks
```


Analysis/Findings...  Both males and females reported a significant decrease in potential to abuse opioids when employed at least part-time. No "full-time" employment data was available from the set. Males especially see significant relief from the potential to abuse via the traditional structure of employment. Females have a slightly higher risk of abuse than males when unemployed - although both males and females do not see a severe increase in abuse potential by being out of work, either. This data seems to illustrate that there is a potentially more valuable connection for males to "work" then females, perhaps due to traditional gender roles and expectations. 


***


## Visualization #3 - The Connection Between Recent Heroine Use and Marital Status
Hypothesis/Question Presented - How often do married people use heroine? What about single people? Divorced? Individuals who use heroine often begin with gateway drugs such as prescription opioids. Does marital data give us any insight based on individual heroine use within the last year?

```{r warning=FALSE}
painrl_new_final %>% #sets up pipe operation
  ggplot(aes(x = HEROINUSE)) + #calls in ggplot function with x-axis
  geom_bar(fill = "hotpink", color = "hotpink")+ #overlays bar graph effect + bar colors
  facet_wrap(~MARRIED, ncol = 1)+ #splits the data into separate visualizations based on marital status with a 1 column limit for a unique view
  coord_flip()+ #flips the x and y data so that the longer text strings can breathe
  theme_economist()+ #adds a eye-catching, editorial-style theme
  labs(title = "The Connection Between Recent Heroine Use and Marital Status", x = "") #changes the titles of variable. In this case I left off the X value to give the x-axis text more room
```

Analysis/Findings...  this was a very interesting visualization. I did not get answers to my original hypothesis/question, but have a new one instead. My new questions is: why do divorced people report so highly on *not* using heroine within the 1 year time frame? Or is it a skew/bias in data due to a higih number of divorced people in the dataset? On a positive note, the data shows us that there is no significantly high group (marital status aside) that has used heroine in the last year.


***
  

## Visualization #4 - Substance Use Treatment Received By Age Groups Within < 1 Year
Hypothesis/Question Presented - 

```{r warning=FALSE}
painrl_new_final %>% #pipe operator setup
  ggplot(aes(x = TRTMENT, fill = AGECAT)) + #sets up ggplot aesthetics
  geom_histogram(position = "dodge", binwidth = 22)+ #overlays histogram graph with thicker binwidth
  labs(title="Substance Use Treatment Received By Age Groups Within < 1 Year", subtitle = "2016-2017", xlab("Number of Individuals"), fill = "Age Bracket")+ #labels all x, y, and title values
  theme_solarized_2()+ #creates a "warm" light effect on the graph for a gentler feel
  theme(axis.title.x=element_blank(), #removes x-axis title
        axis.text.x=element_blank(), #removes x-axis text
        axis.ticks.x=element_blank()) #removes x-axis ticks
```

Analysis/Findings...  The 18-25 age bracket sought treatment the most for substance use disorder between 2016 and 2017 according to this recently available data. 26 years old and up all sought treatment to a much lesser degree and this slowly ticks down in numbers over the lifespan. This is "good" on one hand - this age group is still learning how to be a true adult and would benefit from treatment at this time. It is also "bad", though, in that this could be a sign of a higher rate of abuse overall in the 18-25 age bracket, as well as a lack of outreach, funding, or support for those who may need treatment in older age brackets. It may also be that older adults decrease their use of substances due to life, health, and other events.


***


## Visualization #5 - Title (Map Style)
Hypothesis/Question Presented - Which states present the highest number of individual deaths in the year 2022 related to opioid overdose?

```{r warning=FALSE}
# install.packages("viridis") #commented out, but this installs a special color palette library for use with maps

library(tidyverse) #loads library for pipe operations
library(ggplot2) #loads library for visualization aesthetics
library(maps) #loads the "maps" library for US "map_data" info
library(viridis) #loads the viridis color palette library

drug_od_map <- read_csv("data-table.csv") #loads my secondary data set for mapping

# Prepare the data by converting state abbreviations to lowercase full names
state_full_names <- tibble( #making a tibble with the STATE abbreviations being converted into full state names as a new variable.
  STATE = state.abb, #abbrev. variable
  full_name = tolower(state.name) #new full_name variables
)

drug_od_map <- drug_od_map %>% #pipe operator setup
  left_join(state_full_names, by = "STATE") #joining on the full_names with the STATE abbreviations for our visualization

us_states <- map_data("state") #pulls in the pre-built map data from the "maps" library!


data_map <- us_states %>% #pipe operator setup
  left_join(drug_od_map, by = c("region" = "full_name")) #merging my drug_od_map data with the maps library set

ggplot(data_map, aes(x = long, y = lat, group = group, fill = DEATHS)) + #establishing the ggplot and aesthetics for a map style plot (lat, long, etc.)
  geom_polygon(color = "white") + #this seems to be the "standard" for mapping with geom polygon
  scale_fill_viridis_c(name = "Annual Deaths") + #our pre-built color palette from viridis library
  labs(title = "Drug Overdose Mortality by State in 2022", #title graph label
       subtitle = "National Center for Health Statistics") #subtitle graph label
```


Analysis/Findings...  
I ended up changing the "fill" of the color palette to "deaths" instead of "rate". I found that the deaths as a number had a greater impact on myself as a viewer and also created a greater spectrum of colors. The only limitation is that rate gives us a percentage of individuals per population. In this case, California and Texas stand out as having the highest number of deaths due to a higher population. Florida, however, is highlighted here as an outlier - they have a smaller population but higher rate of annual drug overdose deaths. This tells me that Florida has a bigger overdose mortality problem compared to Texas. Additional research may tell us that Texas and California have greater use of Narcan on overdose patients at the time of opioid abuse.

Source for this analysis: https://www.britannica.com/topic/largest-U-S-state-by-population

***


